\section*{Abstract}

Comparative analyses can provide novel insight into marine population dynamics and the status of fished species, but the world's main stock assessment database (the Myers Stock-Recruitment database) is now outdated.  
To facilitate new analyses, we developed a new stock assessment database, the RAM Legacy database, for commercially exploited marine fishes and invertebrates. 
Time series of total biomass, spawner biomass, recruits, fishing mortality, and catch form the core of the database. 
Assessments were assembled from REF:SQL:TOTNUMMGMT national and international management agencies for a total of REF:SQL:TOTNUMASSESSMENT stocks (REF:SQL:TOTNUMASSESSFISH fish stocks representing REF:SQL:TOTNUMFISHFAMILIES families, and REF:SQL:TOTNUMASSESSINVERT invertebrate stocks representing REF:SQL:TOTNUMINVERTFAMILIES families), including 9 of the world's 10 largest fisheries. Stock assessments were available from REF:SQL:TOTNUMLMESPRIMARY Large Marine Ecosystems and 4 High Seas regions, and include the Atlantic, Pacific, Indian, Arctic and Antarctic Oceans. 
Most assessments came from the U.S., Europe, Canada, New Zealand, and Australia. Assessed marine stocks represent a small proportion of harvested fish taxa (REF:SQL:SRDBPERCENTSPPSAUP\%), and an even smaller proportion of marine fish biodiversity (REF:SQL:SRDBPERCENTSPPFISHBASE\%), but provide high quality abundance data for these intensively studied stocks. 
The database provides new insight into the status of exploited populations: REF:SQL:PERCENTASSESSMENTSBELOWBMSY\% of stocks with reference points (n=REF:SQL:NUMASSESSFRIEDEGG) were estimated to be below the biomass that results in maximum sustainable yield ( ), and REF:SQL:PERCENTASSESSMENTSABOVEFMSY\% had exploitation levels estimated to be above the exploitation rate that results in maximum sustainable yield ( ).
We anticipate that the database will facilitate new research in population and fishery dynamics and life histories and encourage further data contributions from stock assessment scientists.

%Assessments were assembled for REF:SQL:TOTNUMASSESSMENT stocks
%(REF:SQL:TOTNUMASSESSFISH fish populations representing
%REF:SQL:TOTNUMFISHFAMILIES families, and REF:SQL:TOTNUMASSESSINVERT
%invertebrate populations representing REF:SQL:TOTNUMINVERTFAMILIES
%families). Assessments were obtained from REF:SQL:TOTNUMMGMT national
%and international management institutions. Stocks
%present in the database come from REF:SQL:TOTNUMLMESPRIMARY Large
%Marine Ecosystems. Assessed marine fish stocks
%comprise a relatively small proportion of harvested taxa
%(REF:SQL:SRDBPERCENTSPPSAUP\%), and an even smaller proportion of
%marine fish biodiversity (REF:SQL:SRDBPERCENTSPPFISHBASE\%). Reference
%points were available or could be calculated for about
%REF:SQL:PERCENTASSESSFRIEDEGG\% of these stocks. The available data
%provide new insight into the status of exploited populations,
%REF:SQL:PERCENTASSESSMENTSBELOWBMSY\% of stocks with reference points
%were estimated to be below $B_{msy}$, and
%REF:SQL:PERCENTASSESSMENTSABOVEFMSY\% had exploitation levels
%estimated to be above $U_{msy}$. Temporal coverage of assessments is
%recent with 90\% of catch time-series ranging REF:SQL:CATCHTIMESPAN90
%and 90\% of biomass time-series ranging REF:SQL:SSBTIMESPAN90.
