%\newpage
\section*{Abstract}

%Data used to assess the status of individual fish stocks range from
%very little information on many of the world's artisanal fisheries, to
%commercial landings, research surveys, and sophisticated population
%dynamics models that integrate many sources of information.  Previous
%evaluations of the state of global fisheries have used catch data,
%which may be poor proxies for fish stock abundances. A global
%compilation of stock assessment data in the mid-1990s enabled
%substantial syntheses of stock status; however its focus was on
%stock-recruitment relationships and it is now 15 years out of date. 

To facilitate global analyses of population dynamics and the status of
fished species, we have assembled a new database, the RAM Legacy
Database, of the most intensively studied commercially exploited
marine fish stocks. Results from assessment models, including time
series of total biomass, spawner biomass, recruits, fishing mortality,
and catch; reference points; and ancillary information on the life
history, management, and assessment methods for each stock.  Here, we
present the first overview of this database and use it to evaluate the
knowledge-base for assessed marine species.  Assessments were
assembled for 324 stocks
(288 fish species representing
45 families, and 36
invertebrate species representing 12
families), including 8 of the world's 10 largest fisheries.
Assessments were obtained from 18 national and international
management institutions, with most coming from North America, Europe,
Australia, New Zealand and the high seas. Stocks present in the
database come from 31 Large Marine Ecosystems
and cover the Atlantic, Pacific, Indian, Mediterranean, Arctic and Antarctic
Ocean. Reference points were available or could be calculated for
about 74\% of these stocks. The available
data provide new insight into the status of exploited populations,
57\% of stocks with reference points
were estimated to be below $B_{msy}$, and
29\% had exploitation levels
estimated to be above $U_{msy}$.  Assessed marine fish stocks comprise
a relatively small proportion of harvested taxa (24\%), and an even
smaller proportion of marine fish biodiversity (1\%). We hope that
access to the database will facilitate new research into life
histories, population dynamics and the effects of fishing and
encourage further data contributions from stock assessment scientists.

%extractions from the
%database provide new insight into the status of exploited
%populations


%Globally, stock assessments were found
%for 324 stocks (288 species
%of fishes representing 45 families and
%36 species of invertebrates representing
%12 families), from 19
%national and international
%management institutions.

\noindent \textbf{Keywords}: marine fisheries, meta-analysis, population dynamics models, relational database, stock assessment, synthesis.

%with XX\% coming from north temperate regions (North
%Atlantic, North Pacific)
%\noindent Keywords: marine fisheries, meta-analysis, population dynamics models, relational database, stock assessment, synthesis.
%\newpage

%  Geographic differences in assessment
%methods show that Statistical Catch at Age (SCA) models are widely
%used by the west coast of the U.S. (XX percent of assessments),
%regional fishery management organizations in the Pacific (XX percent
%of assessments), and New Zealand (XX percent of assessments); the east
%coast of the U.S. is transitioning from Virtual Population Analysis
%(VPA) to SCA (XX percent of assessments conducted since 2000 have used
%SCA); while VPA is still the dominant assessment
%technique in western Europe (XX percent of assessments).
