%\newpage
%\section*{Abstract}

Data used to assess the status of individual fish stocks varies from
very little information on many of the world's artisanal fisheries, to
commercial landings at various levels of geographic and taxonomic
aggregation, research surveys, and sophisticated population dynamics
models that integrate many sources of information.  Previous
evaluations of the state of global fisheries have used catch or
landings data, which may be poor proxies for fish stock abundances. A
global compilation of stock assessment data in the mid-1990s enabled
substantial syntheses of stock status; however its focus was on
stock-recruitment relationships and it is now 15 years out of date. To
facilitate contemporary syntheses, we have assembled a database of the
most intensively studied commercially exploited marine fish stocks,
including time series of: total biomass, spawner biomass, recruits,
fishing mortality, and catch; reference points; and ancillary
information on the life history, management, and assessment methods
for each stock.  Here, we present the first overview of this database
and use it to evaluate the knowledge-base for assessed marine fishes.
Globally, stock assessments were found for REF:SQL:TOTNUMASSESSMENT
stocks (REF:SQL:TOTNUMASSESSFISH species of fishes representing
REF:SQL:TOTNUMFISHFAMILIES families and REF:SQL:TOTNUMASSESSINVERT
species of invertebrates representing REF:SQL:TOTNUMINVERTFAMILIES
families), from REF:SQL:TOTNUMMGMT national and international
management institutions. Assessments are available for only XX percent
of global marine fisheries catches by weight and XX percent by value.
There is substantial spatial variation in availability of assessed
stocks, with XX percent coming from north temperate regions (North
Atlantic, North Pacific).  Geographic differences in assessment
methods show that Statistical Catch at Age (SCA) models are widely
used by the west coast of the U.S. (XX percent of assessments),
regional fishery management organizations in the Pacific (XX percent
of assessments), and New Zealand (XX percent of assessments); the east
coast of the U.S. is transitioning from Virtual Population Analysis
(VPA) to SCA (XX percent of assessments conducted since 2000 have used
SCA); while VPA is still the dominant assessment
technique in western Europe (XX percent of assessments).\\

%\noindent Keywords: marine fisheries, meta-analysis, population dynamics models, relational database, stock assessment, synthesis.
%\newpage
