\documentclass[letterpaper,12pt]{letter}

\usepackage[margin=1in]{geometry}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[dvips,pdftex]{graphicx}

\name{Daniel Ricard (ricardd@mathstat.dal.ca)\\Dept. of Biological Sciences\\Dalhousie University\\Halifax, NS, B3H 4J1\\Canada}
%\date{October 2008}

\begin{document}
\begin{letter}
{Professor Tony J. Pitcher\\Peter Wall Institute of Advanced Studies\\6331 Crescent Road\\Vancouver, BC, Canada, V6T 1Z2}

\opening{Dear Dr. Pitcher,}
% justification
Please find attached a manuscript titled ``\textit{Assessing the knowledge-base for commercially exploited marine fishes and invertebrates with a new global database of stock assessments}'', that we wish to submit
for review to Fish and Fisheries. The article was encouraged during pre-submission discussions with yourself and Prof. Paul Hart.

We present the first global compilation of stock assessment data since
the Stock-Recruitment Database of our late colleague Ransom A. Myers,
compiled in the 1990s. We build and extend upon this pioneering work
and believe that the present database will make an important
contribution to fisheries research. In accordance with the response to
our presubmission enquiry, we have have focused on a synthesis of the
database in the form of an assessment of the knowledge base for
exploited marine fish. On a global scale, we report on the geographic,
taxonomic and temporal coverage and availability of assessments. We
compare the taxonomic coverage of caught taxa as a subset of marine
fish taxa and assessed taxa as a subset of caught taxa. Using
MSY-based reference points, the status of assessed species is reported
and shown to exhibit marked regional variation. We apply a suite of
novel graphical techniques throughout such that the multi-dimensional
contents of the database and most pertinent results can be conveyed to
the reader.

We are committed to making the database available to other
practitioners and consider Fish and Fisheries the best journal to
reach our intended audience of fisheries scientists.

We would like to propose the following 3 reviewers for this manuscript:

\begin{enumerate}
 \item \textbf{Dr. Rainer Froese} - Leibniz Institute of Marine Sciences, D\"{u}sternbrooker Weg 20, 24105 Kiel, Germany - Phone: +49 (431) 600 4579, Fax: +49 (431) 600 1699 - Email: rfroese@ifm-geomar.de
 \item \textbf{Dr. Keith Sainsbury} - Division of Marine Research, CSIRO GPO Box 1538, Hobart, Tasmania 7001, Australia - Phone: +61 3 6232 5369, Fax: +61 3 6232 5199 - Email: Keith.Sainsbury@csiro.au
 \item \textbf{Dr. Randall Peterman} - School of Resource and Environmental Management, Simon Fraser University, 8888 University Drive, Burnaby, British Columbia, Canada V5A 1S6 - Phone: (778) 782-4683, Fax: (778) 782-4968 - Email: peterman@sfu.ca
\end{enumerate}

These practitioners conduct research in themes related to the global
status of fish stocks and have not been associated with the present
work. Dr. Froese provided us with a detailed taxonomic list of fish
species from FishBase but was otherwise not involved in any of the
analyses. The material contained in the submitted manuscript is
original and has not been submitted elsewhere.

We thank you for the opportunity to submit this work for consideration
as an article in Fish and Fisheries and look forward to hearing from
you.


\closing{Sincerely,}

\cc{C{\'o}il{\'i}n Minto, Marine and Freshwater Research Centre, Galway-Mayo Institute of Technology, Dublin Road, Galway, Ireland}
\cc{Olaf Jensen, Institute of Marine and Coastal Sciences, Rutgers University, USA}
\cc{Julia Baum, National Center for Ecological Analysis and Synthesis, University of California in Santa Barbara, USA}
\end{letter}
\end{document}
